Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-22 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
more edits in eft sections and appendix

In Fig.~\ref{fig:cEFT_cdf} we show the CDF for both events, calculated in the
same way as already described for the EdGB and dCS theories.
%
We see that curves are very similar to those of dCS gravity for GW200129 (see
bottom panel in Fig.~\ref{fig:cEFT_cdf}).
%
Moreover, we find that the EFT~\eqref{eq:eft_bound} and
ParSpec~\eqref{eq:parspec_bound} bounds are satisfied for both events both when
$\mathfrak{m} = M_{\rm f}$ and $\varepsilon = 1/2$.
%
This allows us to place the combined bound of
%
\begin{equation}
    \ell_{\rm cEFT} \leqslant 38.2~\textrm{km} \quad \textrm{at 90\% credible level}\,.
\end{equation}
%
As also happened for our study for dCS, the find that, for
the cubic EFT, the ParSpec bound is violated by the $N_{\rm max} = 1$ corrections to
the QNMs, meaning that we cannot use this case to draw any meaningful
constraint on this parameter.
%
We summarize our results in Table~\ref{tab:summary_ceft}.

@@

@@

\begin{table}[b]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914  & Yes & Yes &  $\ell_{\rm cEFT} \leqslant 38.2$~km \\
0 & GW200129  & Yes & Yes &  $\ell_{\rm cEFT} \leqslant 42.5$~km \\
  & Combined  &     &     &  \cellcolor{black!10}$\ell_{\rm cEFT} \leqslant 38.2$~km \\
\hline
  & GW150914  & Yes  & No  &  -- \\
1 & GW200129  & Yes  & No  &  -- \\
  & Combined  &      &     &  -- \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the cubic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{\rm f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
}
\label{tab:summary_ceft}
\end{table}

@@

For the nonrotating QNMs we used the numerical data of Ref.~\cite{Blazquez-Salcedo:2016enn} and
generated a new quartic fit in $\gamma_{\rm EdGB} = \ell_{\rm EdCS} / M$ using numerical QNM data valid for small
values of the coupling constant [see, in particular, Eq.~(27) and Fig.~2 of Ref.~\cite{Blazquez-Salcedo:2016enn}].
%
We find,
%
\begin{subequations}
\begin{align}
    M {\rm Re}(\sigma)_{\rm EdGB} &= M {\rm Re}(\sigma)_{\rm GR} \, \left( 1 + 0.0107 \, \gamma_{\rm EdGB}^4 \right),
    \\
    M {\rm Im}(\sigma)_{\rm EdGB} &= M {\rm Im}(\sigma)_{\rm GR} \, \left( 1 - 0.0044 \, \gamma_{\rm EdGB}^4 \right).
\end{align}
\end{subequations}
%
The small values of the numerical prefactors of $\gamma_{\rm EdGB}^4$ are a consequence of the how weakly the QNMs
of BHs in EdGB gravity deviate from their GR counterparts, even at moderate values $(\ell_{\rm EdGB} / M)^4 \approx 0.3$.

The rotating corrections are included using the results of Ref.~\cite{Pierini:2021jxd}, in particular their Eqs.~(51) and~(52),
which for consistency with our previous discussion, we truncated at quartic-order in $\gamma_{\rm EdGB}$.
%
Taken together, use as our fitting expressions for the QNM is EdGB gravity
%
\begin{subequations}
\begin{align}
    M {\rm Re}(\sigma)_{\rm EdGB} &= {\rm Re}(\omega)_{\rm GR} \, \left[ 1 + 0.0107 (\ell_{\rm EdGB} / M)^4 \right],
    \\
    M {\rm Im}(\sigma)_{\rm EdGB} &= \dots % {\rm Im}(\omega)_{\rm GR} \, \left[ 1 - 0.0044 (\ell_{\rm EdGB} / M)^4 \right].
\end{align}
\end{subequations}

\dots

>>>>

@@

In Fig.~\ref{fig:cEFT_cdf} we show the CDF for both events, calculated in the
same way as already described for the EdGB and dCS theories.
%
We see that curves are very similar to those of dCS gravity for GW200129 (see
bottom panel in Fig.~\ref{fig:cEFT_cdf}).
%
Moreover, we find that the EFT~\eqref{eq:eft_bound} and
ParSpec~\eqref{eq:parspec_bound} bounds are satisfied for both events both when
$\mathfrak{m} = M_{\rm f}$ and $\varepsilon = 1/2$.
%
This allows us to place the combined bound of
%
\begin{equation}
    \ell_{\rm cEFT} \leqslant 38.2~\textrm{km}, \quad \textrm{at 90\% credible level}\,.
\end{equation}
%
As also happened for our study for dCS, the find that, for
the cubic EFT, the ParSpec bound is violated by the $N_{\rm max} = 1$ corrections to
the QNMs, meaning that we cannot use this case to draw any meaningful
constraint on this parameter.
%
We summarize our results in Table~\ref{tab:summary_ceft}.

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914  & Yes & Yes &  $\ell_{\rm cEFT} \leqslant 38.2$~km \\
0 & GW200129  & Yes & Yes &  $\ell_{\rm cEFT} \leqslant 42.5$~km \\
  & Combined  &     &     &  \cellcolor{black!10}$\ell_{\rm cEFT} \leqslant 38.2$~km \\
\hline
  & GW150914  & Yes  & No  &  -- \\
1 & GW200129  & Yes  & No  &  -- \\
  & Combined  &      &     &  -- \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the cubic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{\rm f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
}
\label{tab:summary_ceft}
\end{table}

@@

For the nonrotating QNMs we used the numerical data of Ref.~\cite{Blazquez-Salcedo:2016enn} and
generated a new quartic fit in $\gamma_{\rm EdGB} = \ell_{\rm EdCS} / M$ using numerical QNM data valid for small
values of the coupling constant [see, in particular, Eq.~(27) and Fig.~2 of Ref.~\cite{Blazquez-Salcedo:2016enn}].
%
We find,
%
\begin{subequations}
\begin{align}
    M {\rm Re}(\sigma)_{\rm EdGB} &= M {\rm Re}(\sigma)_{\rm GR} \, \left( 1 + 0.0107 \, \gamma_{\rm EdGB}^4 \right),
    \\
    M {\rm Im}(\sigma)_{\rm EdGB} &= M {\rm Im}(\sigma)_{\rm GR} \, \left( 1 - 0.0044 \, \gamma_{\rm EdGB}^4 \right).
\end{align}
\end{subequations}
%
The small values of the numerical prefactors of $\gamma_{\rm EdGB}^4$ are a consequence of the how weakly the QNMs
of BHs in EdGB gravity deviate from their GR counterparts, even at moderate values $\gamma_{\rm EdGB} \approx 0.3$.

The rotating corrections are included using the results of Ref.~\cite{Pierini:2021jxd}, in particular their Eqs.~(51) and~(52),
which for consistency with our previous discussion, we truncated at quartic-order in $\gamma_{\rm EdGB}$.
%
The resulting formula can then be expanded in $\gamma_{\rm EdGB}$ and the coefficients $\delta \omega^{(i)}$ and $\delta \tau^{(i)}$, $i=1,2$
read-off by comparison against Eqs.~\eqref{eq:qnm_nongr_iso}.
%
These steps yield,
%
\begin{align}
    p_{\rm EdGB} = 4, \quad \delta \omega_{\rm EdGB}^{(0)} = -0.0107, \quad \delta \tau_{\rm EdGB}^{(0)} = -0.2480,
    \nonumber \\
    \delta \omega_{\rm EdGB}^{(1)} = -0.2480, \quad \delta \tau_{\rm EdGB}^{(1)} = -1.101.
\end{align}

%
% Taken together, use as our fitting expressions for the QNM is EdGB gravity
% %
% \begin{subequations}
% \begin{align}
%     M {\rm Re}(\sigma)_{\rm EdGB} &= {\rm Re}(\omega)_{\rm GR} \, \left[ 1 + 0.0107 (\ell_{\rm EdGB} / M)^4 \right],
%     \\
%     M {\rm Im}(\sigma)_{\rm EdGB} &= \dots % {\rm Im}(\omega)_{\rm GR} \, \left[ 1 - 0.0044 (\ell_{\rm EdGB} / M)^4 \right].
% \end{align}
% \end{subequations}
%
% \dots
